A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-08-03 Analysis/models.Rmd
add scatter and R^2, update text to reflect R's comments

data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  filter(after_mistake==0) %>% 
  select(subject, word_num, word, rt, sentence, type)

@@

tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
          `Log Lik`=map_chr(model_str, log_like)) %>% 
   select(-model_str) %>% write_rds(here("Analysis/maze_model_compare.rds")) %>%  knitr::kable()

@@

d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000)
```

@@

do_comparison <- function(m1, m2){
  # is there benefit of m2 above m1
  if (m1==m2) {return("")}
  lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr)
  lm2 <- lm(str_c(no_surp,m1,m2), data=item_mean_spr)
  a <- anova(lm1,lm2)
  fval <- anova(lm1,lm2)$F[2] %>% round() %>%  as.character
  pval <- anova(lm1,lm2)$`Pr(>F)`[2] %>% pvalue(add_p=T) %>% as.character()
  return(str_c(fval," (",pval,")"))
}

log_like <- function(m1){
    lm(str_c(no_surp,m1), data=item_mean_spr) %>% logLik() %>% round() %>% as.character()

@@

tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
          `Log Lik`=map_chr(model_str, log_like)) %>% 
   select(-model_str)%>% write_rds(here("Analysis/spr_model_compare.rds")) %>% kable()

@@

>>>>

data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  filter(after_mistake==0) %>% 
  select(subject, word_num, word, rt, sentence, type) %>% write_rds(here("Data/maze_pre_error.rds"))

@@

r_squared <- function(m1){
    lm1 <- lm(str_c(no_surp,m1), data=item_mean_data) %>% summary()
    lm1$r.squared %>% round(2) %>% as.character()
}

tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
          `Log Lik`=map_chr(model_str, log_like),
          r_squared=map_chr(model_str, r_squared)) %>% 
   select(-model_str) %>% write_rds(here("Analysis/maze_model_compare.rds")) %>%  knitr::kable()

@@

d.good <- d.limited %>% filter(correct>4) %>% rename(rt=RT) %>% filter(rt>100 & rt<5000) %>% write_rds(here("Data/SPR/first.rds"))

@@

do_comparison <- function(m1, m2){
  # is there benefit of m2 above m1
  if (m1==m2) {return("")}
  lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr)
  lm2 <- lm(str_c(no_surp,m1,m2), data=item_mean_spr)
  a <- anova(lm1,lm2)
  fval <- anova(lm1,lm2)$F[2] %>% round() %>%  as.character
  pval <- anova(lm1,lm2)$`Pr(>F)`[2] %>% pvalue(add_p=T) %>% as.character()
  return(str_c(fval," (",pval,")"))
  #return(a)
}

#do_comparison(ngram, grnn)
log_like <- function(m1){
    lm(str_c(no_surp,m1), data=item_mean_spr) %>% logLik() %>% round() %>% as.character()

@@

r_squared <- function(m1){
    lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr) %>% summary()
    lm1$r.squared %>% round(2) %>% as.character()
}

tibble(Model=c("5-gram","GRNN", "TXL", "GPT-2"), model_str=c(ngram,grnn,txl,gpt),`over 5-gram`=ngram,
             `over GRNN`=grnn,`over TXL`=txl,`over GPT-2`=gpt) %>% 
   mutate(across(`over 5-gram`:`over GPT-2`, ~map2_chr(.x, model_str,do_comparison)),
          `Log Lik`=map_chr(model_str, log_like),
          `r_squared`=map_chr(model_str, r_squared)) %>% 
   select(-model_str)%>% write_rds(here("Analysis/spr_model_compare.rds")) %>% kable()

@@

# R-squared

```{r}
spr <- read_rds(here("Analysis/spr_model_compare.rds"))

maze <- read_rds(here("Analysis/maze_model_compare.rds"))
```

# Maze and SPR scatter

```{r}
tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
  read_delim(delim="\t") %>% 
  mutate(Word_In_Story_Num=zone,
         Story_Num=item)

labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

spr <- read_rds(here("Data/SPR/first.rds")) %>% left_join(tokenization) %>% left_join(labs) %>% select(WorkerId,rt,Story_Num,Sentence_Num, word_num)

maze <- read_rds(here("Data/maze_pre_error.rds")) %>% left_join(labs) %>% select(subject, rt, Story_Num, Sentence_Num, word_num)
```

Abnormally low and high times were excluded, then for each word, an average was taken across subjects. For sentence level, these averages were then summed across the sentence. (Because of A-maze sentence discontinuation and general outlier handling, summing the average is better than averaging the sum.)

Word level correlation between maze and spr
```{r}

summ_spr <- spr %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt))

summ <- maze %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% left_join(summ_spr)

ggplot(summ, aes(x=maze, y=spr))+
  geom_point(alpha=.02)+
  coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
  labs(title="Word level correlation in ms",subtitle="Axes trimmed")

cor(summ$spr,summ$maze, use="complete.obs")

sentence level correlation between maze and spr

Note that a lot of this is driven purely by sentence length, which isn't interesting. 
```{r}
sentence <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=sum(maze)/1000,
            spr=sum(spr)/1000)

ggplot(sentence, aes(x=maze, y=spr))+geom_point(alpha=.1)+
  labs(title="Sentence level correlation in s")

cor(sentence$spr,sentence$maze, use="complete.obs")
```

Averaging across words at the sentence level removes the sentence-length effect.

```{r}
sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=mean(maze),
            spr=mean(spr))
ggplot(sentence_avg, aes(x=maze, y=spr))+geom_point(alpha=.1)+
    coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
  labs(title="Sentence level correlation in ms")

cor(sentence_avg$spr,sentence_avg$maze, use="complete.obs")

As a baseline, we can do all the same things on split-halfs of each data set. We split the subjects in half. (Note that this means that each half has the same # of subjects, but may not have the same #s of each story)

Within spr correlations are higher than spr-maze.

```{r spr-split-half}
set.seed(42)
spr_subs <- spr$WorkerId %>% unique()
length(spr_subs)
#180 participants

spr1 <- sample(spr_subs, 90)

spr_split <- spr %>% 
  mutate(half=ifelse(WorkerId %in% spr1, "spr1","spr2")) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=spr)

ggplot(spr_split, aes(x=spr1, y=spr2))+
  geom_point(alpha=.02)+
  coord_cartesian(xlim=c(0,600),ylim=c(0,600))+
  labs(title="Word level correlation in ms",subtitle="Axes trimmed")

cor(spr_split$spr1,spr_split$spr2, use="complete.obs")
spr_sentence <- spr_split %>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(spr1=sum(spr1)/1000,
            spr2=sum(spr2)/1000)

ggplot(spr_sentence, aes(x=spr1, y=spr2))+geom_point(alpha=.1)+
  labs(title="Sentence level correlation in s")

cor(spr_sentence$spr1,spr_sentence$spr2, use="complete.obs")

spr_sentence_avg <- spr_split%>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(spr1=mean(spr1),
            spr2=mean(spr2))
ggplot(spr_sentence_avg, aes(x=spr1, y=spr2))+geom_point(alpha=.1)+
    coord_cartesian(xlim=c(0,600),ylim=c(0,600))+
  labs(title="Sentence level correlation in ms")

cor(spr_sentence_avg$spr1,spr_sentence_avg$spr2, use="complete.obs")
```

Same things for maze. Note that with fewer subjects, and given only pre-error data, there's a lot more chance of emptiness. I think the weak sentence average correlation is from places where some sentences didn't go as long in one of the halves (b/c low data).

```{r maze-split-half}
set.seed(42)
maze_subs <- maze$subject %>% unique()
length(maze_subs)
#180 participants

maze1 <- sample(maze_subs, 31)

maze_split <- maze %>% 
  mutate(half=ifelse(subject %in% maze1, "maze1","maze2")) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=maze)

ggplot(maze_split, aes(x=maze1, y=maze2))+
  geom_point(alpha=.02)+
  coord_cartesian(xlim=c(0,1500),ylim=c(0,1500))+
  labs(title="Word level correlation in ms",subtitle="Axes trimmed")

cor(maze_split$maze1,maze_split$maze2, use="complete.obs")
maze_sentence <- maze_split %>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze1=sum(maze1)/1000,
            maze2=sum(maze2)/1000)

ggplot(maze_sentence, aes(x=maze1, y=maze2))+geom_point(alpha=.1)+
  labs(title="Sentence level correlation in s")

cor(maze_sentence$maze1,maze_sentence$maze2, use="complete.obs")

maze_sentence_avg <- maze_split%>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze1=mean(maze1),
            maze2=mean(maze2))
ggplot(maze_sentence_avg, aes(x=maze1, y=maze2))+geom_point(alpha=.1)+
    coord_cartesian(xlim=c(0,1500),ylim=c(0,1500))+
  labs(title="Sentence level correlation in ms")

cor(maze_sentence_avg$maze1,maze_sentence_avg$maze2, use="complete.obs")
```

### 2021-08-03 Papers/Paper/manuscript.Rmd
add scatter and R^2, update text to reflect R's comments

abstract: |
  We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.  How long it takes to read a word in a sentence is reflective of how hard it is to identify and integrate the word in the surrounding context. Techniques that slow down the reading process and localize the processing time for each word are useful to understanding the time course of language processing. A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities. We adapt A-maze for use on longer passages and test it on the Natural Stories corpus. We find that people can comprehend what they read during the Maze task. Moreover, the  task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word. This expands the types of effects that can be studied with A-maze, showing it to be a a versatile alternative to eye-tracking and self-paced reading.

@@

One effect of this lack of localization is that reading time for a word is dependent not only on how unexpected a word is, but also how unexpected the previous word is. This is an indication of spillover from the previous word. It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020]. Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization. In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive [@klieglLengthFrequencyPredictability2004].

An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@freedman85; @forsterMazeTaskMeasuring2009]. In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not. Participants must choose the correct word, and their reaction time (RT) is the dependent measure. If participants make a mistake, the sentence discontinues. Theoretically, participants must fully integrate each word into the sentence in order to confidently select it. This idea is supported by studies finding strongly localized effects [@witzelComparisonsOnlineReading2012a].

The downside of Maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task was not widely adopted. @boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence. The quality of these A-Maze distractors is not up to that of hand-generated distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a. A-maze out performed a SPR control in detecting and localizing expected slowdown effects. @sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.

@@

Using this tweak, we test A-maze on the passages from the Natural Stories corpus. The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers. At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions. Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants. If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on basically any naturalistic text. The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension. @futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to. Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles showed a linear relationship to the surprisal of the words.

@@

## Materials
We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], split them into sentences, and ran the sentences through the A-maze generation process.
We follow the A-maze generation process outlined in @boyceMazeMadeEasy2020, although we use an updated version of the codebase with fixes some issues identified in that paper (code at https://github.com/vboyce/Maze). Additionally, the capability to appropriately handle a wider variety of punctuation (needed for this corpus) was added. We took the auto-generated distractors as they were, without checking them for quality. 
We used the original comprehension questions provided in the Natural Stories corpus. To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions. All materials are available at TODO NEW REPO.

@@

To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.  For length, we used the length in characters excluding end punctuation. For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances. We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies. We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020]. For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.

We only included words that were treated as a single token by every model and for which we had frequency information. This mainly excluded words with punctuation as well as uncommon or proper nouns. We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted). We exclude words where mistakes occurred or which occurred after a mistake in the same sentence. We only analysed words where we had values for all predictors.

For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability. We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions. We collapse the reading times across subjects by modelling the mean RT for each word.

@@

In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran. They recruited 181 participants, most of whom read 5 of the stories. After reading each story, each participant answered 6 binary-choice comprehension questions. As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020]. For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct. To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models. 
For linear models, we centered all predictors. We were unable to fit the full mixed effects model. The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and  a per-word-token random intercept, fit with lme4, as this structure did not fit reliably in brms.

@@

Another check is whether participants comprehended the story. We counted how many of the binary-choice comprehension questions each participant got right (out of 6). As seen in Figure \@ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions. Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet. We can't rule out that some participants may have been able to guess the answers without understanding the story. Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task. The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read. Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads. Note that these differences cannot be directly attributed to methods, as the participant populations differed.

@@

Given that the GAM models show a roughly linear relationship, we fit mixed linear models to quantify the influence of surprisal, frequency and word length. 
We built linear models with surprisal, frequency, length and surprisal x length and frequency x length effects from the current and previous words as predictors.

@@

Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare). In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. Overall, F-values here are much lower than for Maze.

@@

This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much [@wilcoxPredictivePowerNeural2020]. 
It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more. Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].

While the error-correction paradigm is crucial to running long materials, it also provides some benefits even on shorter materials when using A-maze with variable participant populations. The error-correction compensates for some of the shortcomings of A-Maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration. Whether this also solves the data loss problem from the researcher perspective within a sentence depends on whether post-mistake data are high-quality and trustworthy; this is a hard-to-assess question of potential interest.

@@

>>>>

abstract: |
  We find support for the localization of reading time effects during the Maze task, as well as extending the range of materials Maze is suitable for.  How long it takes to read a word in a sentence is reflective of how hard it is to identify and integrate the word in the surrounding context. Techniques that slow down the reading process and localize the processing time for each word are useful to understanding the time course of language processing. A-maze is a new method for measuring incremental sentence processing that can localize slowdowns related to syntactic ambiguities. We adapt A-maze for use on longer passages and test it on the Natural Stories corpus. We find that people can comprehend these longer text passages during the Maze task. Moreover, the  task yields useable reaction time data with word predictability effects that are linear in the surprisal of the current word, with little spillover effect from the surprisal of the previous word. This expands the types of effects that can be studied with A-maze, showing it to be a a versatile alternative to eye-tracking and self-paced reading.

@@

One effect of this lack of localization is that reading time for a word is dependent not only on how unexpected a word is, but also how unexpected the previous word is. This is an indication of spillover from the previous word. It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural2020; @goodkindPredictivePowerWord2018; @lukeLimitsLexicalPrediction2016]. Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization. In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive [@klieglLengthFrequencyPredictability2004].

An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@freedman85; @forsterMazeTaskMeasuring2009]. In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not. Participants must choose the correct word, and their reaction time (RT) is the dependent measure. If participants make a mistake, the sentence discontinues. Two versions of the maze task exist: lexical or L-maze where the distractors are nonce words and grammatical or G-maze where the distractors are real words of the language that don't fit in the context of the sentence so far. Theoretically, participants must fully integrate each word into the sentence in order to confidently select it. This idea is supported by studies finding strongly localized effects for G-maze which is more sensitive than L-maze [@witzelComparisonsOnlineReading2012a].

The downside of G-maze is that materials are effort-intensive to construct because of the need to select infelicitious words as distractors for each spot of each sentence; this may explain why the Maze task was not widely adopted. @boyceMazeMadeEasy2020 demonstrate a way to automatically generate Maze distractors by using NLP language models to find words that are high-surprisal in the context of the target sentence. The quality of these A-maze distractors is not up to that of hand-generated G-maze distractors, but @boyceMazeMadeEasy2020 found that materials with A-maze distractors had similar results to the hand-generated distractors from @witzelComparisonsOnlineReading2012a. A-maze outperformed L-maze and an SPR control in detecting and localizing expected slowdown effects. @sloggettAmazeAnyOther2020 also found that A-maze and G-maze distractors yielded similar results on a disambiguation paradigm.

@@

Using this tweak, we test A-maze on the passages from the Natural Stories corpus. The Natural Stories corpus [@futrellNaturalStoriesCorpus2020] consists of 10 passages each roughly 1000 words long which are designed to read fluently to native speakers. At the same time, the passages contain copious punctuation, quoted speech, many proper nouns, and low frequency grammatical constructions. Taken together, these properties make this a severe test of our process, as these features make it harder for the language model to choose good distractors and require focus from participants. If participants can succeed at the Maze task on this set of materials, we think they are likely to succeed on a wide variety of naturalistic texts. The corpus is accompanied by binary-choice comprehension questions, 6 per story, which we use to assess comprehension. @futrellNaturalStoriesCorpus2020 collected self-paced reading time data on this corpus, which we compare the A-maze results to. Participants were able to read and understand these long passages using A-maze with the error correction paradigm, and their RT profiles showed a linear relationship to the surprisal of the words.

@@

## Materials
We took the texts of the Natural Stories corpus [@futrellNaturalStoriesCorpus2020], split them into sentences, and ran the sentences through the A-maze generation process.
We follow the A-maze generation process outlined in @boyceMazeMadeEasy2020, although we use an updated version of the codebase with fixes some issues identified in that paper (code at https://github.com/vboyce/Maze). Additionally, the capability to appropriately handle a wider variety of punctuation (needed for this corpus) was added. We took the auto-generated distractors as they were, without checking them for quality. 
We used the original comprehension questions provided in the Natural Stories corpus. To familiarize participants with the task, we wrote a short practice passage and corresponding comprehension questions. See the Appendix for an except of one of the stories and its corresponding comprehension questions. All materials are available at TODO NEW REPO.

@@

```{r}
surps <- read_rds(here("Prep_code/natural_stories_surprisals.rds"))

not_first <- surps %>% filter(Word_In_Sentence_Num!=1)

#nrow(not_first)

not_na <- not_first %>% filter(across(ends_with("surp"),~!is.na(.x)),
                               !is.na(freq))

not_ngram <- not_na %>%   filter(ngram_token_count==1)

#nrow(not_ngram)

not_other <- not_ngram %>% 
filter(txl_token_count==1)%>% 
  filter(grnn_token_count==1)
#nrow(not_other)

not_gpt <- not_other %>% filter(gpt_token_count==1)

#nrow(not_gpt)

```
To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.  For length, we used the length in characters excluding end punctuation. For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances. We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies. We got per-word surprisals for each of 4 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], Transformer-XL [@daiTransformerXLAttentiveLanguage2019], and GPT-2 [@radfordLanguageModelsAre], using lm-zoo [@gauthierSyntaxGymOnlinePlatform2020].

We exclude the first word of every sentence because it has a different distractor, which leaves `r nrow(not_first)` words. Then we exclude words for which we don't have surprisal or frequency information, leaving `r nrow(not_na)` words. We additionally exclude words that were treated as being composed of multiple tokens. This is primarily words that have punctuation, where the punctuation is treated as a separate token. While surprisals should be additive, the summing the surprisals over these parts gives unreasonable responses sometimes. For instance, in one story the word king!' is given a surprisal of 64 by GRNN (context: The other birds gave out one by one and when the eagle saw this he thought, 'What is the use of flying any higher? This victory is in the bag and I am king!'). To avoid these outliers we exclude all words that a model treated as multi-token, leaving `r nrow(not_gpt)` words. (While GPT-2 using byte-pair encoding that can split up words into multiple parts, excluding words it split up only excludes 30 words that were not already excluded by other models.)

For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word. We excluded outlier RTs that were <100 or >5000 ms (<100 is likely a recording error, >5000 is likely the participant getting distracted). We exclude words where mistakes occurred or which occurred after a mistake in the same sentence. We only analysed words where we had values for all predictors, which means that if the previous word was unknown to a model, this word will be excluded because we're missing values for a lagged predictor.

For generalized additive models, we centered but did not rescale the length and frequency predictors, but left surprisal uncentered for interpretability. We used smooths for the surprisal terms and tensor effects for the frequency by length effects and interactions. To minimize the effect of repeated measures on confidence about inferred curve shapes, we collapse the reading times across subjects by modelling the mean RT for each word.

@@

In addition to the texts, @futrellNaturalStoriesCorpus2020 released reading time data from a SPR study they ran in 2011. They recruited 181 participants from Amazon Mechanical Turk, most of whom read 5 of the stories. After reading each story, each participant answered 6 binary-choice comprehension questions. As a comparison to our A-maze models, we run similar models on the SPR corpus on Natural Stories [@futrellNaturalStoriesCorpus2020]. For comparability, we analyse only the first story each participant read, and, in line with @futrellNaturalStoriesCorpus2020, exclude participants who got less than 5/6 of the comprehension questions correct. To account for spill over effects known to exist in SPR, we analyse predictors at the current word as well as the past 3 words for all models. 
For linear models, we centered all predictors. We were unable to fit the full mixed effects model. The best model we could fit had by-subject random intercept, uncorrelated by-subject random slopes for surprisal, length and frequency, and  a per-word-token random intercept, fit with lme4, as this structure did not fit reliably in brms.

@@

Another check is whether participants comprehended the story. We counted how many of the binary-choice comprehension questions each participant got right (out of 6). As seen in Figure \@ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions. Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet. We can't rule out that some participants may have been able to guess the answers without understanding the story. Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task. The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read. Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads. Note that these differences cannot be directly attributed to methods, as the participant populations differed. While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected.

@@

Given that the GAM models show a roughly linear relationship, we fit mixed linear models to quantify the influence of surprisal, frequency and word length. 
We built linear models with surprisal, frequency, length and surprisal x length and frequency x length effects from the current and previous words as predictors. We centered, but did not rescale, the predictors.

@@

Similarly to the model comparison for Maze, we also conducted a model comparison on the SPR data shown in Table \@ref(tab:spr-compare). In terms of log likelihoods, we find that GPT-2 is better than 5-gram is better than GRNN is better than TXL, although differences are small. Model comparisons show that GPT-2 and 5-gram models contain some value over each other model, which is less clear for TXL and GRNN. The relatively good fit of 5-gram models to SPR data compared with neural models matches results from @huSystematicAssessmentSyntactic2020 and @wilcoxPredictivePowerNeural2020. This contrasts with the Maze results, where the 5-gram model has the worst fit and does not provide additional predictive value to the other models.

```{r}
maze <- read_rds(here("Analysis/maze_model_compare.rds"))
spr <- read_rds(here("Analysis/spr_model_compare.rds"))
```

As an overall measure of fit to data, we calculate multiple R-squared for the single surprisal source models for both A-maze and SPR. The models predict A-maze better with R-squared values ranging from `r maze %>% filter(Model=="5-gram") %>% pull(r_squared)` for the 5-gram model to `r maze %>% filter(Model=="GPT-2") %>% pull(r_squared)` for GPT-2. Whereas for SPR, the R-squared values range from from `r spr %>% filter(Model=="5-gram") %>% pull(r_squared)` to `r spr %>% filter(Model=="GPT-2") %>% pull(r_squared)`. This suggests that the effect size differences are not due merely to the larger overall reading time for A-maze, but that instead A-maze is more sensitive to surprisal and length effects.

@@

Actually the potential applicability and value is much broader, for any question where nailing down the locus on incremental processing difficulty is important, little spillover is very valuable.

This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty and thus benefits from the lack of spillover. 
It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more. Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].

While the error-correction paradigm is crucial to running long materials, it also provides some benefits even on shorter materials when using A-maze with variable participant populations. The error-correction compensates for some of the shortcomings of A-maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration. One questions that remains is whether this post-mistake data is useable, which depends on whether RTs from a few words after a mistake show any differences from RTs before any mistakes. Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is a  hard-to-assess question of potential interest.

@@

# Appendix 
The beginning of one of the stories. This is the first 200 words of a 1000 word story.

Tulip mania was a period in the Dutch Golden Age during which contract prices for bulbs of the recently introduced tulip reached extraordinarily high levels and then suddenly collapsed. At the peak of tulip mania in February sixteen thirty-seven, tulip contracts sold for more than ten times the annual income of a skilled craftsman. It is generally considered the first recorded economic bubble. The tulip, introduced to Europe in the mid sixteenth century from the Ottoman Empire, became very popular in the United Provinces, which we now know as the Netherlands. Tulip cultivation in the United Provinces is generally thought to have started in earnest around fifteen ninety-three, after the Flemish botanist Charles de l'Ecluse had taken up a post at the University of Leiden and established a botanical garden, which is famous as one of the oldest in the world. There, he planted his collection of tulip bulbs that the Emperor's ambassador sent to him from Turkey, which were able to tolerate the harsher conditions of the northern climate. It was shortly thereafter that the tulips began to grow in popularity. The flower rapidly became a coveted luxury item and a status symbol, and a profusion of varieties followed.

The first 2 out of the 6 comprehension questions.

When did tulip mania reach its peak? 1630's, 1730's

From which country did tulips come to Europe? Turkey, Egypt
